### 2025-08-01 data-visualization.qmd
20250801 - tweaks

The `ggplot2` package [@R-ggplot2] is based on the grammar of graphics [@ggplot3rdEd], which is the principle that you can build any graph from three components: 1) the data set, 2) the coordinate system, and 3) the geoms—visual marks that represent data points.
The general syntax for plots created using the `ggplot2` package [@R-ggplot2] is:

@@

ggplot2::ggplot(
  data = INSERT, # 1) data set
  mapping = aes(
    x = variableName1, # name of variable on x-axis
    y = variableName2) # name of variable on y-axis
) +
  coord_cartesian( # 2) coordinate system
    xlim = c(NA, NA), # set limits (min and max) of x-axis
    ylim = c(NA, NA)) + # set limits (min and max) of y-axis
  INSERT_GEOM_FUNCTIONS # 3) geom(s)
```

@@

ggplot2::ggplot(
  data = INSERT, # 1) data set
  mapping = aes(
    x = variableName1, # name of variable on x-axis
    y = variableName2) # name of variable on y-axis
) +
  coord_cartesian( # 2) coordinate system
    xlim = c(NA, NA), # set limits (min and max) of x-axis
    ylim = c(NA, NA)) + # set limits (min and max) of y-axis
  INSERT_GEOM_FUNCTIONS + # 3) geom(s)
  scale_x_continuous(
    breaks = INSERT # specify x-axis labels
  ) +
  scale_y_continuous(
    breaks = INSERT # specify y-axis labels
  ) +
  labs( # add plot labels
    x = "INSERT", # x-axis title
    y = "INSERT", # y-axis title
    title = "INSERT", # plot title
    subtitle = "INSERT", # plot subtitle
    caption = "Add a caption below plot", # plot caption
    alt = "Add alt text to the plot" # plot alt-text
  ) +
  INSERT_THEME_NAME + # apply theme
  facet_wrap(vars(INSERT_VARIABLES_TO_FACET)) # facet by one or more variables
```

### Geoms {#sec-ggplot2SyntaxGeoms}

Here is a list of common geoms available in `ggplot2` (there are many others):

@@

ggplot2::ggplot(
  data = rb_seasonal,
  aes(
    x = age,
    y = rushing_yards)) +
  geom_point() +
  geom_smooth() +
  geom_rug(
    col = "#800000",
    alpha = 0.2) +
  coord_cartesian(
    xlim = c(20,40),
    ylim = c(0,NA)) +
  scale_x_continuous(
    breaks = seq(from = 20, to = 40, by = 5)
  ) +
  scale_y_continuous(
    breaks = seq(from = 0, to = 2500, by = 250)
  ) +
  labs(
    x = "Running Back's Age (years)",
    y = "Rushing Yards (Season)",
    title = "NFL Rushing Yards (Season) by Player Age",
    subtitle = "(Among Running Backs)"
  ) +
  theme_classic()
```

>>>>

The `ggplot2` package [@R-ggplot2] is based on the grammar of graphics [@ggplot3rdEd], which is the principle that you can build any graph from three components: 1) the data, 2) the coordinate system and aesthetics, and 3) the geometries ("geoms")—visual marks that represent data points.
The general syntax for plots created using the `ggplot2` package [@R-ggplot2] is:

@@

ggplot2::ggplot(
  data = INSERT, # 1) data
  mapping = aes(
    x = variableName1, # name of variable on x-axis
    y = variableName2) # name of variable on y-axis
) +
  coord_cartesian( # 2) coordinate system
    xlim = c(NA, NA), # set limits (min and max) of x-axis
    ylim = c(NA, NA)) + # set limits (min and max) of y-axis
  INSERT_GEOM_FUNCTIONS # 3) geom(s)
```

@@

ggplot2::ggplot(
  data = INSERT, # 1) data
  mapping = aes(
    x = variableName1, # name of variable on x-axis
    y = variableName2) # name of variable on y-axis
) +
  coord_cartesian( # 2) coordinate system
    xlim = c(NA, NA), # set limits (min and max) of x-axis
    ylim = c(NA, NA)) + # set limits (min and max) of y-axis
  INSERT_GEOM_FUNCTIONS + # 3) geom(s)
  scale_x_continuous(
    breaks = INSERT # specify x-axis labels
  ) +
  scale_y_continuous(
    breaks = INSERT # specify y-axis labels
  ) +
  labs( # add plot labels
    x = "INSERT", # x-axis title
    y = "INSERT", # y-axis title
    title = "INSERT", # plot title
    subtitle = "INSERT", # plot subtitle
    caption = "Add a caption below plot", # plot caption
    alt = "Add alt text to the plot" # plot alt-text
  ) +
  INSERT_THEME_NAME + # apply theme
  facet_wrap(vars(INSERT_VARIABLES_TO_FACET)) # facet by one or more variables
```

### Geometries {#sec-ggplot2SyntaxGeoms}

Here is a list of common geometrics ("geoms") available in `ggplot2` (there are many others):

@@

ggplot2::ggplot(
  data = rb_seasonal,
  aes(
    x = age,
    y = rushing_yards)) +
  geom_point() +
  geom_smooth() +
  geom_rug(
    color = "#800000",
    alpha = 0.2) +
  coord_cartesian(
    xlim = c(20,40),
    ylim = c(0,NA)) +
  scale_x_continuous(
    breaks = seq(from = 20, to = 40, by = 5)
  ) +
  scale_y_continuous(
    breaks = seq(from = 0, to = 2500, by = 250)
  ) +
  labs(
    x = "Running Back's Age (years)",
    y = "Rushing Yards (Season)",
    title = "NFL Rushing Yards (Season) by Player Age",
    subtitle = "(Among Running Backs)"
  ) +
  theme_classic()
```
